Responses to Reviewer Three

### 2021-05-11 Scripts/Manuscript.Rmd
Ended first revision

Our findings highlighting early shifts in the shared distribution of transboundary stocks concur with previous studies that have detected and attributed changes in marine catch composition [e.g., @Cheung:2013gk; @Frainer:2017ji; @Last:2011fd]. For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (US) as a response to climatic, oceanographic and ecological changes [@Zeidberg:2007gs]. A new fishery targeting Humboldt squid quickly developed on the heels of the species’ range expansion [@Pinsky:2014by]. In the northeast Atlantic, fisheries on Atlantic mackerel (*Scomber scombrus*) are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC). However, a range expansion of Atlantic mackerel into Icelandic waters in 2007 resulted in Iceland capturing 6% of the fishery’s total allowable catch and a further 18% in 2008, without consultation with NEAFC, threatening the sustainability of the stock [@Spijkers:2017ij]. These changes resulted in disputes between Iceland and the Faroe Islands, as well as among NEAFC member states [@Spijkers:2017ij]. As transboundary stocks continue to shift, such international disputes are likely to increase in frequency and requiring a revision of management plans [@Pinsky:2018cb; @Engler:2020bg].

@@

**Management plans that are not designed to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp]. Thus, focusing on these “hotspots” is key to anticipate a potential increase in fisheries conflicts in coming years. Identified strategies to cope with changes in the shared proportion of transboundary stocks include strengthening of current cooperative mechanisms and the consideration of side payments, which do not have to be monetary [@Tunca:2019if; @Miller:2013iv], increased international cooperation [@Miller:2013iv] and management rules that capture distributional shifts [@PalaciosAbrantes:2020vt; @Pinsky:2018cb]. Some examples of current treaties that have adapted some of these strategies include the Pacific Salmon Treaty between Canada and the U.S. to manage Pacific salmon (Oncorhynchus sp.) that has a conservation fund that works as a side payment [@Miller:2013iv], the International Pacific Halibut Commission between Canada and the United States that manages Pacific halibut (*Hippoglossus stenolepis*) and allocates quota based on the stock’s yearly distribution [@PalaciosAbrantes:2020vt], and the Parties to the Nauru Agreement (PNA) in the Pacific Island States that manage skipjack tuna (*Katsuwonus pelamis*). The PNA manage their fisheries under a so-called Vessel Day Scheme, which accounts for shifts in stocks and catches, and is therefore somewhat anticipatory [@Aqorau:2018bh]. Quota allocation methods based on a stock’s current distribution and/or on fixed-historical proportion like that of the European Union need to evolve to be more agile [@Baudron:2020dj] and potentially move towards a dynamic method or a combination of both [@PalaciosAbrantes:2020vt; @Sumaila:2020tp]. However, even management plans that take into consideration such strategies might not be fully prepared for the consequences of shifting transboundary stocks [@Pinsky:2018cb; @Koubrak:2020hw; @Engler:2020bg]. For example species’ shifts within the PNA area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb]. Moreover, the transition from historic to dynamic allocations can find strong resistance from stakeholders “losing” benefits from a fishery they have historically been entitled. Finally, changes in policy can take years (e.g., it took the PST about 10 years to agree on a conservation fund) and sometimes, the time of political change is longer than that of species shifts [@Pinsky:2012kq].**

>>>>

Our findings highlighting early shifts in the shared distribution of transboundary stocks concur with previous studies that have detected and attributed changes in marine catch composition [e.g., @Cheung:2013gk; @Frainer:2017ji; @Last:2011fd]. For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (US) as a response to climatic, oceanographic and ecological changes [@Zeidberg:2007gs]. A new fishery targeting Humboldt squid quickly developed on the heels of the species’ range expansion [@Pinsky:2014by]. In the northeast Atlantic, fisheries on Atlantic mackerel (*Scomber scombrus*) are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC). However, a range expansion of Atlantic mackerel into Icelandic waters in 2007 resulted in Iceland capturing 6% of the fishery’s total allowable catch and a further 18% in 2008, without consultation with NEAFC, threatening the sustainability of the stock [@Spijkers:2017ij]. These changes resulted in disputes between Iceland and the Faroe Islands, as well as among NEAFC member states [@Spijkers:2017ij]. **Early range expansions across international jurisdictions have also been documented for multiple stocks along the European Union regulatory areas [@Baudron:2020dj], the Benguela Current [@Potts:2014bq; @Yemane:2014jg], and the southwest South Atlantic Ocean [@Franco:2020bk].**As transboundary stocks continue to shift, international disputes are likely to increase in frequency and stocks' health decrease requiring international collaboration and revision of management plans [@Pinsky:2018cb; @Engler:2020bg; @Maureaud:2021gj].

@@

**Management plans that are not designed to respond to range and abundance shifts will be less resilient to climate change [@Miller:2013iv; @Sumaila:2020tp]. Thus, focusing on these “hotspots” is key to anticipate a potential increase in fisheries conflicts in coming years. Identified strategies to cope with changes in the shared proportion of transboundary stocks include strengthening of current cooperative mechanisms and the consideration of side payments, which do not have to be monetary [@Tunca:2019if; @Miller:2013iv], increased international cooperation [@Miller:2013iv] and management rules that capture distributional shifts [@PalaciosAbrantes:2020vt; @Pinsky:2018cb]. Some examples of current treaties that have adapted some of these strategies include the Pacific Salmon Treaty between Canada and the U.S. to manage Pacific salmon (Oncorhynchus sp.) that has a conservation fund that works as a side payment [@Miller:2013iv], the International Pacific Halibut Commission between Canada and the United States that manages Pacific halibut (*Hippoglossus stenolepis*) and allocates quota based on the stock’s yearly distribution [@PalaciosAbrantes:2020vt], and the Parties to the Nauru Agreement (PNA) in the Pacific Island States that manage skipjack tuna (*Katsuwonus pelamis*). The PNA manage their fisheries under a so-called Vessel Day Scheme, which accounts for shifts in stocks and catches, and is therefore somewhat anticipatory [@Aqorau:2018bh]. Quota allocation methods based on a stock’s current distribution and/or on fixed-historical proportion like that of the European Union need to evolve to be more agile [@Baudron:2020dj] and potentially move towards a dynamic method or a combination of both [@PalaciosAbrantes:2020vt; @Sumaila:2020tp]. However, even management plans that take into consideration such strategies might not be fully prepared for the consequences of shifting transboundary stocks [@Pinsky:2018cb; @Koubrak:2020hw; @Engler:2020bg]. For example species’ shifts within the PNA area will likely also have to deal with the ‘newcomer issue’ that NEAFC is currently facing regarding Atlantic mackerel [@Spijkers:2017ij; @Pinsky:2018cb]. Moreover, the transition from historic to dynamic allocations can find strong resistance from stakeholders “losing” benefits from a fishery they have historically been entitled. Finally, changes in policy can take years (e.g., it took the PST about 10 years to agree on a conservation fund) and sometimes, the time of political change is longer than that of species shifts [@Pinsky:2012kq]. This is specifically true as in many cases we lack standardized data to track shifts of species across international jurisdictions [@Maureaud:2021gj] as these might already be happening.**

### 2021-05-11 Scripts/Reviews/reviewer_one.Rmd
Ended first revision

 - **Regarding the bi-lateral agreements**, this is a great question that, unfortunately, we cannot respond at this moment. Currently, there is only one database on bi-lateral agreements produce by Oremus *et al* (2020). On their paper, the authors state "we analysed 127 publicly available international fisheries agreements in the South Pacific, Southeast Asia, South Asia, Northwest Africa, and Central America and the Caribbean" and " For each agreement, we recorded the following information... If relevant, to which specific stock or species the agreement pertained". We contacted Dr. (Kimberly) Oremus who provided the provided the mentioned dataset. However, non of the 127 treaties have identified subjects at the species level (except for bigeye, skipjack and yellowtail tunas).

We hope that the reviewer understands the complexity of building such database and that this exercise falls outside the scope and stage of our current analysis. We agree that having such information would increase the manuscript impact, but, would not necessarily directly impact our results nor the message we want to covein (Ultimately Oremus *et at* state it very clearly in their paper, "None of the 127 agreements contains language that is directly related to climate change, range shifts or stock exits").

@@

 - **Regarding model-specific variance.** For this paper we used one climate model (the GFDL) forced following one climate change scenario (RCP 8.5) to couple with one species distribution model. For this reason, providing quantitative upper and lower bounds of our results is not possible and would require substantial extra computation. However, because we have 10 different runs of the GFDL and employ the time of emergence concept, we are capable of providing a less conservative estimate of the numbers you mention (i.e., the number of shifting stocks and EEZs). This was originally presented in the methods section as a sensitivity analysis and has now been moved to the Methods section of the main paper (**NOTE TO CO-AUTHORS**: Should we include this in the results or leave it in the methods?). It reads on line *XXX* of page *XX*: *We also computed a less conservative estimation of shifting transboundary stocks by including a threshold of one standard deviation (i.e., representing a probability of 68% that the stock had shifted from natural variation; see methods). In this case, the global average shift in the shared distribution of transboundary stocks is 2029 $\pm$ 27 years, a difference of about 7 years (Fig. S8). In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks (Figs. S2 and S8).*

@@

- **Response:** Yes, "transboundary between neighboring countries” is indeed redundant. We have modified the manuscript to say "*Climate change is shifting the distribution of shared fish stocks between neighboring countries (transboundary).*" We believe it is important to keep both parts of the sentence for clarification.

@@

- **Response:** Similarly, we do not have space to discuss this here. Please note that we discuss management plans on the paragraph starting on line *13* of page *12*.

@@

- **Response:** We have changed this sentence to *Marine species’ distributions mainly reflect species’ preferences for discrete environmental conditions (2).*

@@

- **Response:** In the section "High present-day climate risk of on transboundary fisheries management" we describe early shifts in the shared distribution of transboundary stocks including the cases of Humboldt squid (*Dosidicus gigas*) and Atlantic mackerel (*Scomber scombrus*) both in the early 2000s and have included a couple more punctual examples of documented shifts between 2006 and 2021. However, there is not a vast documentation on shifts in the distribution of transboundary species (i.e., specifically from one EEZ to another). Mapping these shifts is out of the scope of the current project as there is an ongoing effort to document the global hifts of transboundary species.

@@

- **Response:** Thank you for your comment. We provide direction of change in the next figure (Figure 5 of the revised manuscript) showing winners and losers according to neighboring EEZs. Here we show how each neighboring EEZ will stand to win or lose against its neighbors. **NOTE FOR CO-AUTHORS:** *I am hesitant of actually pointing out winners and losers (e.g., stating that Spain will loos most of its catch to France). Hence, I rather leave the results and discussion more general. Thoughts?*

@@

- **Response:** We have modified this sentence in the revised manuscript with the hopes to make it less jargon-y. It now reeds on line *7* of *P9* the revised manuscript: "*There will be an increase in the number of stocks changing their catch proportion and the number of EEZs with changing stocks towards the mid-21st century. Such increase will be followed by an also increase in the intensity of the proportion change (Figs. S3 and S4).*"

@@

- **Response:** Thank you for this comment. We have edited the paragraph starting at line **13** of page **12** to reflect this comment.

@@

- **Response:** Thank you for highlighting this. We have changed the sentence to reflect this comment.

>>>>

 - **Regarding the bi-lateral agreements**, this is a great question that, unfortunately, we cannot respond at this moment. Currently, there is only one database on bi/multi-lateral agreements produced by Oremus *et al* (2020). On their paper, the authors "analysed 127 publicly available international fisheries agreements in the South Pacific, Southeast Asia, South Asia, Northwest Africa, and Central America and the Caribbean" and "For each agreement, we [the authors] recorded the following information... If relevant, to which specific stock or species the agreement pertained". We contacted Dr. (Kimberly) Oremus who provided  the mentioned data-set. However, non of the 127 treaties have identified subjects at the species level (except for bigeye, skipjack and yellowtail tunas). Thus, we were unable to match the stocks with existing agreements.

We hope that the reviewer understands the complexity of building such database and that this exercise falls outside the scope and stage of our current analysis. We agree that having such information would increase the manuscript impact, but, would not necessarily affect our results, nor the message we want to covey (Ultimately Oremus *et at* state it very clearly in their paper, "None of the 127 agreements contains language that is directly related to climate change, range shifts or stock exits").

@@

 - **Regarding model-specific variance.** For this paper we used one climate model (the GFDL) forced following one climate change scenario (RCP 8.5) to couple with one species distribution model. For this reason, providing quantitative upper and lower bounds of our results is not possible and would require substantial extra computation. However, because we have 10 different runs of the GFDL and employ the time of emergence concept, we are capable of providing a less conservative estimate of the numbers you mention (i.e., the number of shifting stocks and EEZs). This was originally presented in the Supplemental Material section as a sensitivity analysis and has now been moved to the Methods section of the revised paper (**NOTE TO CO-AUTHORS**: Should we include this in the results or leave it in the methods?). It reads on line *XXX* of page *XX*: *We also computed a less conservative estimation of shifting transboundary stocks by including a threshold of one standard deviation (i.e., representing a probability of 68% that the stock had shifted from natural variation; see methods). In this case, the global average shift in the shared distribution of transboundary stocks is 2029 $\pm$ 27 years, a difference of about 7 years (Fig. S8). In total, 5,745 transboundary stocks would shift their shared distributions between 2006 and 2100 considering a less conservative metric, that is, 33% (n = 1626) more shifting stocks (Figs. S2 and S8).*

@@

- **Response:** Yes, "transboundary between neighboring countries” is indeed redundant, thank you for noticing. We have modified the manuscript to say "*Climate change is shifting the distribution of shared fish stocks between neighboring countries (transboundary).*" We believe it is important to keep both parts of the sentence for clarification.

@@

- **Response:** Similarly, we do not have space to discuss this here. Please note that we discuss management plans on the paragraph starting on line *XXX* of page *XX*.

@@

- **Response:** We have changed this sentence to reflect the suggestion, it now reads: *Marine species’ distributions mainly reflect species’ preferences for discrete environmental conditions (2).*

@@

- **Response:** There is not a vast documentation on shifts in the distribution of transboundary species (i.e., specifically from one EEZ to another). This is partially because, from a scientific perspective, there is not enough standardized data to track changes across countries (See Maureaud *et all*, 2021 below). Mapping these shifts is out of the scope of the current project as there is an ongoing effort to document the global hotspots of transboundary species. That being said, in the section "High present-day climate risk of on transboundary fisheries management" we describe early shifts in the shared distribution of transboundary stocks including the cases of Humboldt squid (*Dosidicus gigas*) and Atlantic mackerel (*Scomber scombrus*) both in the early 2000s. Moreover, we have included a couple of other examples of shifts reported in neighbouring countries. It reads on line **XX** of page **X** of the revised manuscript: *Early range expansions across international jurisdictions have also been documented for multiple stocks along the European Union regulatory areas [@Baudron:2020dj], the Benguela Current [@Potts:2014bq; @Yemane:2014jg], and the southwest South Atlantic Ocean [@Franco:2020bk].*

@@

- **Response:** Thank you for your comment. We provide direction of change in the next figure (Figure 5 of the revised manuscript) showing winners and losers according to neighboring EEZs. This analysis is based on the time of emergence result, that has no direction, we show how each neighboring EEZ will stand to win or lose against its neighbors.

@@

- **Response:** Thank you for noticing this. We have modified this sentence in the revised manuscript with the hopes to make it less jargon-y. It now reeds on line *XX* of *XXX* the revised manuscript: "*There will be an increase in the number of stocks changing their catch proportion and the number of EEZs with changing stocks towards the mid-21st century. Such increase will be followed by an also increase in the intensity of the proportion change (Figs. S3 and S4).*"

@@

- **Response:** Thank you for this comment. We have edited the paragraph starting at line **XXX** of page **XX** to reflect this comment.

@@

- **Response:** We agre and have changed the sentence to reflect this comment.

### 2021-05-11 Scripts/Reviews/reviewer_three.Rmd
Ended first revision

- **Response:** We appreciate the reviewer's comment regarding our method and understand the concern on consider national stocks that become transboundary stocks and *vice versa*, however, these questions have already been responded by Pinsky *et al* (2018) and Oremus *et al.* (2020) and consequently, follow outside the scope of the current paper.

In their 2018 paper, Pinsky *et al* look at how many *new* transboundary species each EEZ of the world will have towards the end of the century under climate change (i.e., national stocks that become transboundary stocks). Indeed, the authors find that "*many of the world’s EEZs are likely to receive one to 12 new, climate-driven transboundary stocks by the end of the century*." On the other hand, Oremus *et al* (2020) "*estimates the number of species exiting national jurisdictions under climate change up to 2100*" concluding that *tropical nations stand to lose the most species under these scenarios.* However, both studies fail to identify changes in the *current* transboundary stock, gap that we look to fill int his paper. Here, we used the species (n = 633) identified by Palacios-Abrantes *et al* (2020) as transboundary and consequently do not have national stocks identified. This was on purpose as the objective of our paper is precisely to look at the impacts of climate change on *current* transboundary species. Addressing the question of entering/exciting stocks would require us to change our initial species list, change our otherwise already complex and long methodology, and likely give the paper a different perspective.

@@

 - **Response:** We meant that 59% is the average proportion of the catch that shifts between countries, or more specific, the stocks reponsable for that catch. We have revised the manuscript to clarify this, it reads on line **XX** of page *X*: "*Findings show that by 2030, 59% $\pm$ 17% of the yearly catch from transboundary stocks is projected to change beyond an EEZ’s historic proportion (i.e., between 2004 and 2010; Fig. 4A)*"

@@

- **Response:** The main structure was following the journal's guideline. We have now included a summarized version of the methods in the main article.

@@

- **Response:** Thank you, we have adopted the suggested approach. You will read in line 14 of page 1: "*Using global climate-fisheries model simulations, we show that 78% the world’s international boundaries will experience a shifting transboundary stock by 2030*".

@@

- **Response:** Thank you for noticing this. We have changed the wording to reflect this comment.

@@

- **Response:** We have included one of the examples we referenced, it now reads in line *XX* of page *xxx in the revised manuscript:* *"For example, Pacific salmon (Oncorhynchus sp.) jointly managed by Canada and the U.S. was heavily fished in the early 1990s after a stock shift favored the Alaskan fleet over the Canadian fleet (18).*"

>>>>

- **Response:** We appreciate the reviewer's comment regarding our method and understand the concern on consider national stocks that become transboundary stocks and *vice versa*, however, these questions follow outside the scope of the current paper as they have already been responded by Pinsky *et al* (2018) and Oremus *et al.* (2020).

In their 2018 paper, Pinsky *et al* look at how many *new* transboundary species each EEZ of the world will have towards the end of the century under climate change (i.e., national stocks that become transboundary stocks). Indeed, the authors find that "*many of the world’s EEZs are likely to receive one to 12 new, climate-driven transboundary stocks by the end of the century*." On the other hand, Oremus *et al* (2020) "*estimates the number of species exiting national jurisdictions under climate change up to 2100*" concluding that *tropical nations stand to lose the most species under these scenarios.* However, both studies fail to identify changes in *current* transboundary stocks, gap that we look to fill int his paper. Here, we used the species (n = 633) identified by Palacios-Abrantes *et al* (2020) as transboundary and consequently do not have national stocks identified. This was on purpose as the objective of our paper is precisely to look at the impacts of climate change on *current* transboundary species. Addressing the question of entering/exciting stocks would require us to change our initial species list, change our otherwise already dense methodology as you rightfully point out, and likely give the paper a different perspective.

@@

 - **Response:** We meant that 59% is the average proportion of the catch that shifts between countries, or more specific, the stocks responsible for that catch. We have revised the manuscript to clarify this, it reads on line **XX** of page *X*: "*Findings show that by 2030, 59% $\pm$ 17% of the yearly catch from transboundary stocks is projected to change beyond an EEZ’s historic proportion (i.e., between 2004 and 2010; Fig. 4A)*"

@@

- **Response:** The main structure was following the journal's guideline. We have now included a summarized version of the methods in the main article. Please see line **XXX** of page **XX*.*

@@

- **Response:** Thank you, we have adopted the suggested approach. It reads now in line 14 of page 1 of the revised manuscript: "*Using global climate-fisheries model simulations, we show that 78% the world’s international boundaries will experience a shifting transboundary stock by 2030*".

@@

- **Response:** Thank you for noticing this. We have changed the wording to reflect this comment in the revised manuscript.

@@

- **Response:** We have included one of the examples we referenced, it now reads in line *XX* of page *xxx *in the revised manuscript:* *"For example, Pacific salmon (Oncorhynchus sp.) jointly managed by Canada and the U.S. was heavily fished in the early 1990s after a stock shift favored the Alaskan fleet over the Canadian fleet (18).*"
